---
title: "Exploration of wierd values of correlation coefficient"
author: ""
date: ""
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo=FALSE)
```


## Checking whether R functions are correct

## 1. Checking `convert_F_to_t` function

Problematic is study of Perkins et al. (2023) I am first creating objects containing numbers reported in this study e.g. pre_mean etc.

```{r include=FALSE}
# Creating variables
post_mean <-  data$`Tratment Group Post Mean`
post_mean

pre_mean <-  data$`Treatment Group Pre Mean`
pre_mean

M_change <- post_mean - pre_mean
M_change

F_value <- data$F_value
F_value

N <- data$`Treatment Group Sample Size`
N

pre_sd <- data$`Treatment Group Pre SD`
post_sd <-data$`Treatment Group Post SD`
t <- convert_F_to_t(F_value)

change_sd <- calculate_change_sd(M_change, N, t)
change_sd
```

Converting f to t
```{r}
t <- convert_F_to_t(data$F_value)
t
```

I use also another formula reported here: https://matthewbjane.com/blog-posts/blog-post-3.html; to check whether it is correct.

```{r}
t <- M_change / (change_sd/sqrt(N))
t
```

Values are the same so t-value is correct so function: `convert_F_to_t` is working correctly. Note that NAs are OK here. 

## 2. Checking `calculate_change_sd` function

As far as I know, there is no way to check whether this function is working correctly by using some alternative method and than compare results. But the code does seems to be straightforward and looks analogous to mathematical formula: 

 **Standard Deviation of Change Scores**:
\[ S_{\text{change}} = \frac{M_{\text{change}} \cdot \sqrt{N}}{t} \]

```r
# Function to calculate change score standard deviation
calculate_change_sd <- function(M_change, N, t) {
  return((M_change * sqrt(N)) / t)
}
```

## 3. Checking `calculate_final_correlation` function

```{r}
calculate_final_correlation(pre_sd = pre_sd, post_sd = post_sd, change_sd = change_sd)
```
This is weird output, but to be sure about that, we will use the code from the following webside: https://matthewbjane.com/blog-posts/blog-post-3.html. The code from this source is as follows:

```{r echo=TRUE}
(t^2*(data$`Treatment Group Pre SD`^2 + data$`Treatment Group Post SD`^2) - N * M_change^2) / (2*t^2*data$`Treatment Group Pre SD`* data$`Treatment Group Post SD`)
```

Ok, the output looks the same, so the source of the problem lies in the reported data itself. Therefore, the necessary step before analysis is to remove studies with values that do not make sense.





