---
title: "Meta-Analysis of Pre/Post Correlations"
author: "Lukas Novak"
date: "2024-06-04"
output:
  html_document: default
---

```{r echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, echo=FALSE)
```

## Introduction

In this analysis, I calculated the pre/post correlation for treatment and control groups using the given dataset. The correlation is calculated based on the following steps:

1. Calculate the standard deviation of change scores \( SD_{\text{change}} \).
2. Final Correlation \( r \).
3. **F** value was converted to **t** value if necessary

## Source of equations

The only source of equations I found is this: https://matthewbjane.com/blog-posts/blog-post-3.html. From this source, all equaions were obtained. The equations are as follows:

# Equations

1. **Standard Deviation of Change Scores**:
\[ S_{\text{change}} = \frac{M_{\text{change}} \cdot \sqrt{N}}{t} \]

2. **Final Correlation**:
\[ r = \frac{S_{\text{pre}}^2 + S_{\text{post}}^2 - S_{\text{change}}^2}{2 \cdot S_{\text{pre}} \cdot S_{\text{post}}} \]

3. **Convert F to t**:
\[ t = \sqrt{F} \]

```{r echo=FALSE}
# Load necessary packages
library(readxl)
library(meta)
library(dplyr)
library(googlesheets4)
```

```{r message=FALSE, warning=FALSE}
# Path to your downloaded JSON key file
json_key <- ".secreats/elevated-column-421419-ca4f60e60a16.json"

# Authenticate using the service account
gs4_auth(path = json_key)

# Define the Google Sheet URL or Sheet ID
sheet_url <- "https://docs.google.com/spreadsheets/d/1IoeSzBbG3lBfPoa6qIspxMZvpbaiVTVjU9ZA0e5DZRY/edit?usp=sharing"

# Read the data from the Google Sheet
data <- read_sheet(sheet_url)

# Print the data to confirm it was loaded correctly
# print(data)
```

```{r include=FALSE}
# Function to randomly assign F_value and SD_change if missing
assign_random_columns <- function(data) {
  data <- data %>%
    rowwise() %>%
    mutate(
      `F_value` = ifelse(runif(1) > 0.5, runif(1, min = 1, max = 10), NA),
      `SD_change` = ifelse(is.na(`F_value`), runif(1, min = 1, max = 10), NA)
    ) %>%
    ungroup()
  return(data)
}

# Apply the function to the data
if(all(is.na(data %>% select(c("F_value","SD_change"))))) {
  print("There are no data in F_calue or SD_change so I am simulating data for them...")
  data <- assign_random_columns(data)
} else {
  print("Ok, there is a either F_value column or SD_change and thus I am using values from these columns")
}

```

```{r echo=FALSE}
print("I am filtering only those studies, which are having value in etiher F_value or SD_change columns")
data = data %>% 
  filter(if_any(c("F_value","SD_change"), ~!is.na(.)))
```

```{r}
# Function to calculate change score standard deviation
calculate_change_sd <- function(M_change, N, t) {
  return((M_change * sqrt(N)) / t)
}

# Function to calculate final correlation
calculate_final_correlation <- function(pre_sd, post_sd, change_sd) {
  return((pre_sd^2 + post_sd^2 - change_sd^2) / (2 * pre_sd * post_sd))
}

# Function to convert F to t
convert_F_to_t <- function(F) {
  return(sqrt(F))
}
```

```{r}
# Function to calculate correlation for a single study using either F_value or SD_change
calculate_study_correlation <- function(pre_sd, post_sd, pre_mean, post_mean, N, F_value, SD_change) {
  if (!is.na(F_value)) {
    t <- convert_F_to_t(F_value)
    M_change <- post_mean - pre_mean
    change_sd <- calculate_change_sd(M_change, N, t)
  } else if (!is.na(SD_change)) {
    change_sd <- SD_change
  } else {
    stop("Both F_value and SD_change are missing")
  }
  return(calculate_final_correlation(pre_sd, post_sd, change_sd))
}
```

```{r}
# Calculate correlation for treatment group
data$Pre_Post_Correlation_Treatment <- mapply(calculate_study_correlation, 
                                              data$`Treatment Group Pre SD`, 
                                              data$`Treatment Group Post SD`, 
                                              data$`Treatment Group Pre Mean`, 
                                              data$`Tratment Group Post Mean`, 
                                              data$`Treatment Group Sample Size`, 
                                              data$F_value, 
                                              data$SD_change)

# Calculate correlation for control group
data$Pre_Post_Correlation_Control <- mapply(calculate_study_correlation, 
                                            data$`Control Pre SD`, 
                                            data$`Control Post SD`, 
                                            data$`Control Pre Mean`, 
                                            data$`Control Post Mean`, 
                                            data$`Control Sample Size`, 
                                            data$F_value, 
                                            data$SD_change)

```

# Results

```{r}
# Correlation between pre and post in treatment and control groups
data %>% 
  select(Author, Pre_Post_Correlation_Treatment, Pre_Post_Correlation_Control) %>% 
  knitr::kable()
```

Ok, some correlations in output exceeds one, this is weird...I am running some diagnostic scripts to figure this out....

```{r renderChildScript, echo=FALSE}
# Render the external Rmd script
rmarkdown::render("./Exploration_of_wierd_values_of_correlation_coefficient.Rmd", output_file = "Exploration_of_wierd_values_of_correlation_coefficient.html", quiet = TRUE)

# Include the rendered HTML in the main document
html_content <- paste(readLines("./Exploration_of_wierd_values_of_correlation_coefficient.html", warn = FALSE), collapse = "\n")
htmltools::HTML(html_content)
```

```{r eval=FALSE, include=FALSE}
# Meta-analysis for treatment group correlations
meta_treatment <- metacor(
  cor = data$Pre_Post_Correlation_Treatment,
  n = data$`Treatment Group Sample Size`,
  studlab = data$Author,
  sm = "ZCOR"  # Using Fisher's z transformation
)

# Meta-analysis for control group correlations
meta_control <- metacor(
  cor = data$Pre_Post_Correlation_Control,
  n = data$`Control Group Sample Size`,
  studlab = data$Author,
  sm = "ZCOR"  # Using Fisher's z transformation
)
```

```{r eval=FALSE, include=FALSE}
# Print the results
print(meta_treatment)
print(meta_control)
```

```{r eval=FALSE, include=FALSE}
# Optionally, you can also plot the forest plots
forest(meta_treatment, fontsize = 8, spacing = .5)
forest(meta_control, fontsize = 8, spacing = .5)
```
